Bayesian exponential random graph modeling of KWL Burial network

### 2020-12-04 analysis/paper/paper.Rmd
revise the bootstrap paragraphs

pre_E <- burial_phases_sample_size %>% filter(Phase == "pre") %>% pull()
post_E <- burial_phases_sample_size %>% filter(Phase == "post") %>% pull()
chi <- burial_phases_sample_size %>% filter(Phase == "chi") %>% pull()
```

@@

<!-- comparison between two networks using patchwork bootstrap --> 
One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials. To understand if this network size different affects the posterior distributions, we conducted a sensitivity test using vertex bootstrap inference on the networks. This allows a comparison for multiple networks of different sizes [@Chen2019]. Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e. node) of a network to quantify standard errors in the network statistics of interest. This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population. We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for two networks. The result shows that the 95% confidence interval of each network statistic contains the observed network of both periods, which indicates our networks are relatively robust to differences in sample size. This suggests a very small sample size effect on both models. In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters. In general, there are no significant changes in the posterior estimates using different sets of priors. But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation. <!-- re-write wither to justify why vertex bootstrap is the right method to answer this question, or use patchwork bootstrap and justify that-->

@@

<!-- methodological implication-->
In this paper, we used burial network as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation [@Binford1971; @Saxe1970]. We also note that the issue of disconnection between the role of the deceased in burial context and their real life, which can be resolved by examined different evidence out of the burial context [@Chapman2003]. The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices. The adoption of Bayesian inference on ERGM can solve the issues of computational tractability and degeneracy in parameter estimation in ERGM [@Caimo2017]. In addition, a Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations. We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation. It is important to note that although some network statistics of nodal attributes do not highly match our assumption, the Bayesian approach is still an efficient and effective method for evaluating the endogenous network effects. This closely related to the overall structure of burials that can be viewed as a representation of a given society.

>>>>

pre_E <- burial_phases_sample_size %>% filter(Phase == "pre") %>% pull()
post_E <- burial_phases_sample_size %>% filter(Phase == "post") %>% pull()
chi <- burial_phases_sample_size %>% filter(Phase == "chi") %>% pull()
diff_pre_post <- post_E - pre_E
```

@@

<!-- comparison between two networks using patchwork bootstrap --> 
One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials. To understand if this network size different affects the posterior distributions, we conducted a sensitivity test by randomly removing `r diff_pre_post ` nodes from the post-European network to resemble the size of pre-European network with vertex bootstrap. Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e. node) of a network to quantify standard errors in the network statistics of interest. This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population, especially for smaller networks. We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for two networks. The result shows that the 95% confidence interval of each network statistic contains the observed network density and degree, but higher extreme network clustering. This pattern is consistent before and after node removal for the post-European network, and the difference can be still identified between networks from different phases. This indicates our networks are relatively robust to differences in sample size and suggests a very small sample size effect.

Similarly, we can use vertex bootstrap to compare pre- and post-European networks since it allows a statistical testing for multiple networks of different sizes [@Chen2019; @Snijders1999]. We used t test for simulated networks and the results show that there are significant difference in all network statistics from both periods. In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters. In general, there are no significant changes in the posterior estimates using different sets of priors. But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation.

@@

<!-- methodological implication-->
In this paper, we used burial network as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation [@Binford1971; @Saxe1970]. We also note that the issue of disconnection between the role of the deceased in burial context and their real life, which can be resolved by examined different evidence out of the burial context [@Chapman2003]. The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices. The adoption of Bayesian inference on ERGM can solve the issues of computational tractability and degeneracy in parameter estimation in ERGM [@Caimo2017]. In addition, a Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations. We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation. Although some network statistics of covariants do not showing the pattern as we assumed, the Bayesian approach is an efficient and effective method for evaluating the endogenous network effects relative to the overall structure of burials. This sturcutre can be viewed as a representation of a given society.
